---
title: "Jags enforcement model"
author: "Rod"
date: "19/5/2022"
output:
  pdf_document: default
  word_document: default
---
Load Packages
```{r include=FALSE}
library(jagsUI)
library(XLConnect) #### THIS REQUIERES JAVA
library(coda)
library(mcmcplots)
#library(tidyverse)
library(rethinking)
library(Metrics)
#library(caret)
library(performanceEstimation)
library(dplyr)
library(ggplot2)
library(MCMCvis)
library(gam)
library(beepr)
library(performance)
library(coda)
library(rjags)

```

Data
```{r include=FALSE}
#### Our data
Binomial_dataset    <- read.csv("~/Dropbox/Chile Dataset analysis/Data and code/Data/Binomial_dataset.csv")   

# Function for standardizing variable between 0 and 1
zsc.fn <- function(x){(x-min(x))/(max(x)-min(x))}
Binomial_dataset    <-na.omit(Binomial_dataset)

```

THE MODEL
```{r echo=TRUE, warning=TRUE}
modC <- "model {
for (i in 1:N) { 
        n.binomial[i]     ~  dbern(P[i])
        
        P[i]              <- prob[i] * SimVio[i] + 0.00001     ## hack required for Rjags -- otherwise 'incompatible'-error 
        
        SimVio[i]         ~  dbern(p[i])

        logit(p[i])       <- g0     +  g1*Time[i]      +  g2*Enforcers[i]   + a[Actor[i]]  + y[Year[i]]  + r[Region[i]]  
       
        logit(prob[i])    <- g3     +  v[Violation[i]] +  s[Specie[i]]      
  }
  
  #Priors

  #Intercepts
  g0  ~ dunif(-1, 1) 
  g3  ~ dunif(-1, 1) 
  
  #Continuos variables
  g1   ~ dunif(-2.2, 2.2) 
  g2   ~ dunif(-2.2, 2.2)
 
  
  #Categorical variables
  
  #Actor
     a[1] <- 0       
  for(d in 2:n_actors) {
    a[d] ~ dunif(-2.2, 2.2)
  }

#Violation
  v[1] <- 0 
  for(c in 2:n_violations) {   
    v[c] ~ dunif(-2.2, 2.2)
  }

  
#Species
     s[1] <- 0
  for(e in 2:(n_species)) {
    s[e] ~ dunif(-2.2, 2.2)
 }
  

#Region
     r[1] <- 0
  for(d in 2:(n_regions)) {
    r[d] ~ dunif(-2.2, 2.2)
  }
 
  #Year
    y[1] <- 0
  for(f in 2:(n_year)) {
    y[f] ~ dunif(-2.2, 2.2)
  }
 

  R    <- P[]
  Pr   <- p[]
  Prob <- prob[]
  V    <- v[]
  A    <- a[]
  S    <- s[]
  Re   <- r[]
  Y    <- y[]
  
} #end model
"
filename.use <- 'jags0.txt'
writeLines(modC, filename.use)  
filename.save <- paste(modC, 'txt', sep = '.') 

  
```


Chunk for preparing data and RUNNING model

*Warning of Rhat values not being calculated is due to one of categories being set to 0
```{r echo=TRUE, warning=TRUE}
basis.dat <- 
  list( N            = dim(Binomial_dataset)[1], 
        n.binomial   = Binomial_dataset$Binomial.Response,                        ### Response variable

        Time         = zsc.fn(Binomial_dataset$Time),                             ### Time Predictor  
        Enforcers    = zsc.fn(Binomial_dataset$N.enforcers.per.code),             ### Number of enforcers per code
        
        n_violations = length(unique(Binomial_dataset$VIOLATION)),                ### Violation types
        Violation    = (as.numeric(as.factor(Binomial_dataset$VIOLATION))),    
       
        n_actors     = length(unique(Binomial_dataset$ACTOR)),                    ### Actor types
        Actor        = (as.numeric(as.factor(Binomial_dataset$ACTOR))),
 
        n_species    = length(unique(Binomial_dataset$SPECIES)),                  ### Species
        Specie       = (as.numeric(as.factor(Binomial_dataset$SPECIES))), 
        
        n_regions    = length(unique(Binomial_dataset$Region)),                  ### Species
        Region       = (as.numeric(as.factor(Binomial_dataset$Region))),
        
        n_year       = length(unique(Binomial_dataset$YEAR)),                  ### Species
        Year         = (as.numeric(as.factor(Binomial_dataset$YEAR))))        ### Week of the year 




#mod.inits <- function(){  
  list(g0      = runif(1, -1, 0), 
       g1      = runif(1, -1, 0),
       g2      = runif(1, -1, 0),
       g3      = runif(1, -1, 0)
  )  


## PARAMETERS
pms <- c('g0', 'g1', 'g2', 'g3', "Y", "Re","V", "A", "S", "R", "Prob", 'Pr' ) # 

start_time <- Sys.time()
start_time

# FIT MODEL
modC <- jags(data = basis.dat,
              inits = mod.inits, ##
              parameters.to.save = pms, 
              model.file = "Jags0.txt",
              parallel   = T,
              n.cores    = 4,   
              n.chains   = 3,     ### Normal is 3   
              n.adapt    = 80,   ### normal is 100
              n.iter     = 1000,  ### Normal is 3000
              n.burnin   = 300,   ### normal is 500
              n.thin     = 1,
              set.seed(150))

mod.use <- modC
end_time <- Sys.time()
end_time - start_time

#start_time <- Sys.time()
#start_time
#mod.use <- update(mod.use,n.iter = 1000)
#start_time <- Sys.time()
#start_time
#mod.use <- update(mod.use,n.iter = 1000)

#mod.use = (autojags(data = basis.dat,
#             inits = mod.inits,
#              parameters.to.save = pms, 
#              model.file = "Jags0.txt",
#              parallel   = T,
#              n.cores    = 4,   
#              n.chains   = 4,    
#              n.adapt=NULL, iter.increment=1500, n.burnin=500, n.thin=1,
#              save.all.iter=FALSE, modules=c('glm'), factories=NULL, 
#              DIC=TRUE, 
#              store.data=FALSE, codaOnly=FALSE,seed=NULL, 
#              bugs.format=FALSE, Rhat.limit=1.1, max.iter=10000, verbose=TRUE))

beep(sound = 8)
```

Evaluating model performance
```{r echo=TRUE, warning=TRUE, fig.keep='all'}
hist(mod.use$summary[,8], main="Rhat values", breaks=10000)
plot(mod.use)
```


Determining threshold for assessing fit
```{r echo=TRUE, warning=TRUE}
results=matrix(0,100,4)

for (cr in 1:100){
crit= cr/100
predicted_violations = ifelse(mod.use$mean$R >crit, 1,0)
F1=confusionMatrix(as.factor(predicted_violations), as.factor(Binomial_dataset$Binomial.Response), mode = "everything", positive="1")
F11=F1$byClass["F1"]
results[cr,1]=crit
results[cr,2]=F11

}
plot(results[,1],results[,2], ylab="F1", xlab="Threshold")
Most.accurate=(which.max(results[,2]))/100

```


- Full evaluation using the threshold from previous chunk
- Full evaluation using threshold set at 0.5

It shows pretty good F1 values for both 
```{r echo=TRUE, warning=TRUE}
predicted_violations = ifelse(mod.use$mean$R > min(0.5,Most.accurate), 1,0)
F1=confusionMatrix(as.factor(predicted_violations), as.factor(Binomial_dataset$Binomial.Response), mode = "everything", positive="1")
F1


predicted_violations = ifelse(mod.use$mean$R > 0.5, 1,0)
F1.2=confusionMatrix(as.factor(predicted_violations), as.factor(Binomial_dataset$Binomial.Response), mode = "everything", positive="1")
F1.2
```


FIGURES!
Distribution graphs
```{r}
p=density(mod.use$mean$Pr, adjust=4)
mean(mod.use$mean$Pr)
sd(mod.use$mean$Pr)
plot(p, main="Probability of Detecting an Existing Violation", xlab= "Probability")
#abline(v=mean)

prob=density(mod.use$mean$Prob, adjust=3)
mean(mod.use$mean$Prob)
sd(mod.use$mean$Prob)
plot(prob, main="Probability of a Violation", xlab= "Probability")
#abline(v=mean)

P=density(mod.use$mean$R, adjust=3)
mean(mod.use$mean$R)
sd(mod.use$mean$R)
plot(P, main="Overall Probability of Detection When Enforcing", xlab= "Probability")
#abline(v=mean)


Time=density(mod.use$sims.list$g1, adjust=3)
mean(mod.use$sims.list$g1)
sd(mod.use$sims.list$g1)
plot(Time, main="Time Parameter Estimate", xlab= "Log odds ratios")
#abline(v=mean)

Enforcers=density(mod.use$sims.list$g2, adjust=3)
mean(mod.use$sims.list$g2)
sd(mod.use$sims.list$g2)
plot(Enforcers, main="Enforcers Parameter Estimate", xlab= "Log odds ratios")
#abline(v=mean)


Interaction=density(mod.use$sims.list$g4, adjust=3)
mean(mod.use$sims.list$g4)
sd(mod.use$sims.list$g4)
plot(Interaction, main="Interaction Estimate", xlab= "Log odds ratios")

```



```{r fig1, fig.height = 8, fig.width = 8}
##Species

Species_names= c("Anchovy and sardine spp (0.15)", "Common Hake (0.08)", "Kelp spp (0.07)", "Clams spp (0.07)","Unclear (0.07)","Centolla spp (0.06)", "Sea urchin (0.06)","Loco (0.05)", "Southern Hake (0.05)", "Chilean Seabass (0.03)", "Conger eel (0.03)", "Mussell spp (0.03)", "Pomfret (0.02)","Crab spp (0.02)", "Jumbo squid (0.02)","Mackerel (0.02)","Others (0.02)","Shrimp spp (0.01)","Scallop spp (0.01)","Various (0.01)","Gigartina spp (0.01)", "Trophon snail (0.009)","Sword fish (0.007)","Lobster (0.007)","Octopus spp (0.007)","Hoki (0.005)","Salmon spp (0.003)","Limpet spp (0.002)","Prawns spp (0.002)","Skate (0.002)","Shark spp (0.001)")

MCMCplot(mod.use, 
         params = 'S', 
       xlim = c(-2.5, 2.5),
         xlab = 'Parameter estimate (log odds ratios)',
         main = 'Species',
         labels= Species_names,
          sz_labels = 1,
          sz_med = 1,
          sz_thick = 1,
        ci = c(5, 90),
          sz_thin = 1,
          sz_ax = 2,
          sz_main_txt = 1)

```


```{r fig1, fig.height = 8, fig.width = 6}
Regions_names=s= c("Arica y Parinacota (0.026)","Tarapaca (0.019)", "Antofagasta (0.057)", "Atacama (0.029)", "Coquimbo (0.045)", "Valparaiso (0.072)", "Metropolitana (0.062)", "Lib B O'Higgins (0.013)", "Maule (0.038)", "Araucanía (0.008)", "Biobío (0.196)", "Los Ríos (0.060)", "Los Lagos (0.235)", "Aysen (0.070)", "Magallanes (0.055)", "Mobile Unit (0.006)")
       
MCMCplot(mod.use, 
         params = 'Re', 
       xlim = c(-2.5, 2.5),
         xlab = 'Parameter estimate (log odds ratios)',
         main = 'Region',
         labels= Regions_names,
          sz_labels = 1,
          sz_med = 1,
          sz_thick = 1,
        ci = c(5, 90),
          sz_thin = 1,
          sz_ax = 2,
          sz_main_txt = 1)
```


```{r fig1, fig.height = 4, fig.width = 8}
##Violation
Violation_names= c("Access (0.34)","Quota (0.27)", "Ban (0.20)","Minimum Size (0.06)","Legal Origin (0.01)", "Other (0.005)","Fishing Gear (0.001)")


MCMCplot(mod.use, 
         params = 'V', 
         xlim = c(-2.5, 2.5),
         xlab = 'Parameter estimate (log odds ratios)',
         main = 'Violation',
         labels= Violation_names,
          sz_labels = 1,
          sz_med = 1,
          sz_thick = 1,
          sz_thin = 1,
          sz_ax = 2,
          sz_main_txt = 1)

##Actor
Actor_names= c( "Small-scale (0.41)","Transporter (0.16)", "Processing Plants (0.12)", "Industrial Boat (0.08)", "Marketers (0.08)", "No Activity (0.08)","Other Actor (0.03)","Restaurant (0.02)")

MCMCplot(mod.use, 
         params = 'A', 
       xlim = c(-2.5, 2.5),
         xlab = 'Parameter estimate (log odds ratios)',
         main = 'Actor',
         labels= Actor_names,
          sz_labels = 1,
          sz_med = 1,
          sz_thick = 1,
          sz_thin = 1,
          sz_ax = 2,
          ci= c(50,90) , 
          sz_main_txt = 1)


##Year
Year_names= c( "2014 (10414)","2015 (11903)", "2016 (13860)", "2017 (10459)", "2018 (10244)", "2019 (9682)","2020 (6993)")

MCMCplot(mod.use, 
         params = 'Y', 
       xlim = c(-2.5, 2.5),
         xlab = 'Parameter estimate (log odds ratios)',
         main = 'Year',
         labels= Year_names,
          sz_labels = 1,
          sz_med = 1,
          sz_thick = 1,
          sz_thin = 1,
          sz_ax = 2,
          ci= c(50,90) , 
          sz_main_txt = 1)

```


Effort against week estimates
```{r}
Binomial_dataset$Ones = 1
EffortPerWeek       = aggregate(Binomial_dataset$Ones ~ OFFICE + MONTH  , data = Binomial_dataset, FUN = sum, na.rm = TRUE)
ViolationsPerWeek   = aggregate(Binomial_dataset$Binomial.Response ~ OFFICE + MONTH  , data = Binomial_dataset, FUN = sum, na.rm = TRUE)

#EffortPerWeek$YEAR = EffortPerWeek$DATE
EffortPerWeek$DATE       = mod.use$mean$M
EffortPerWeek$Violations = ViolationsPerWeek$`Binomial_dataset$Binomial.Response`
names(EffortPerWeek)[names(EffortPerWeek) == 'Binomial_dataset$Ones'] <- 'Effort'
#names(EffortPerWeek)[names(EffortPerWeek) == 'Binomial_dataset$N.actions.per.code'] <- 'Effort'
#EffortPerWeek <- subset(EffortPerWeek,EffortPerWeek$Effort<300)


###Direct Week
ggplot(EffortPerWeek, aes(x = (Effort), y =DATE) ) + geom_point() + geom_smooth(method = "glm", se = FALSE) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+ xlab(as.expression(expression( paste("Effort per Office per Month") ))) + 
  ylab("Detectability Parameter Estimate")
aa=lm(EffortPerWeek$DATE ~EffortPerWeek$Effort)
summary(aa)


##Violations 
ggplot(EffortPerWeek, aes(x = (Effort), y =Violations) ) + geom_point() + geom_smooth(method = "glm", se = FALSE) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+ xlab(as.expression(expression( paste("Effort per Office per Month") ))) + 
  ylab("Number of visits with violations found")
aa=lm(EffortPerWeek$Violations ~EffortPerWeek$Effort)
summary(aa)

###Effort previous week
Newone = matrix(0,559,2)
Newone[,1] = EffortPerWeek$Effort[1:559]
Newone[,2] = EffortPerWeek$DATE[2:560]
EffortPerWeek = as.data.frame(Newone)
colnames(EffortPerWeek) = c("DATE", "Effort")


ggplot(EffortPerWeek, aes(x = (DATE), y =(Effort)) ) + geom_point() + geom_smooth(method = "glm", se = FALSE) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+ xlab(as.expression(expression( paste("Effort per Week") ))) + 
  ylab("Weekly Detectability Parameter Estimate") 
aa=lm(EffortPerWeek$DATE ~EffortPerWeek$Effort)
summary(aa)

```

Descriptive stats
```{r eval=FALSE, include=FALSE}
Binomial_dataset["One"] = 1
Effort = aggregate(One~SPECIES,FUN=sum, data=Binomial_dataset)
Violations = aggregate(Number.of.sanctions~SPECIES,FUN=sum, data=Binomial_dataset)

Effort["Violations"] = Violations$Number.of.sanctions
Effort$SPECIES = Violation_names


Regions_names=s= c("Arica y Parinacota","Tarapaca", "Antofagasta", "Atacama", "Coquimbo", "Valparaiso", "Metropolitana", "Lib B O'Higgins", "Maule", "Ñuble", "Araucanía", "Biobío", "Los Ríos", "Los Lagos", "Aysen", "Magallanes", "Mobile Unit")


Species_names= c("Anchovy and sardine spp","Centolla spp", "Chilean Seabass", "Clams spp","Common Hake", "Conger eel", "Crab spp"    , "Gigartina spp", "Hoki", "Jumbo squid)", "Kelp spp","Limpet spp", "Lobster","Loco","Mackerel","Mussell spp", "Octopus spp", "Others","Pomfret","Prawns spp",  "Salmon spp","Scallop spp", "Sea urchin","Shark spp", "Shrimp spp", "Skate", "Southern Hake", "Sword fish", "Trophon snail", "Unclear", "Various")

Violation_names= c("Access", "Ban","Quota","MinSize","FishingGear","Other","LegalOrigin")


write.xlsx(Effort,"~/Dropbox/Region.xlsx", row.names = FALSE)


aa = table(Binomial_dataset$SPECIES)
bb = lm(mod.use$mean$S ~as.numeric(aa))
summary(bb)

aa= order(Effort$One)
```

Effort against probability of detection
```{r eval=FALSE, include=FALSE}
data=matrix(0,length((Binomial_dataset$Time)),2)
data[,1]= (Binomial_dataset$Time*Binomial_dataset$N.enforcers.per.code)/60
data[,2]= mod.use$mean$R
data=as.data.frame(data)


aa= ggplot(data, aes(x=V1, y=V2)) + geom_point() +
             geom_smooth(method="gam", se=TRUE) + ylim(0,1) + xlab("Enforcer hours") + ylab("Probability of Detection") + theme_classic()


model   <- loess(V2 ~ lo(V1+2), data = data)
data$prediction <- predict(model)

max=data$V1[which.max(data$prediction)]

aa  + geom_vline(xintercept = max, linetype = 2, color= "red")


```

```{r}
##Actor
Actor_names= c( "Pescador Artesanal (41%)","Transportistas (16%)", "Plantas Procesadoras (12%)", "Pescador Industrial (8%)", "Comercializadoras (8%)", "Sin Actividad (8%)","Otros Actores (8%)","Restaurantes (2%)")

MCMCplot(mod.use, 
         params = 'A', 
       xlim = c(-2.5, 2.5),
         xlab = 'Resultados del Modelo (log odds ratios)',
         main = 'Tipo de Actor',
         labels= Actor_names,
          sz_labels = 1,
          sz_med = 1,
          sz_thick = 1,
          sz_thin = 1,
          sz_ax = 2,
          ci= c(50,90) , 
          sz_main_txt = 1)

```

